CODEBOOK FOR GETTING AND CLEANING DATA PROJECT
========================================================

The Getting and Cleaning Data Coursera Course Project involved preparing a tidy data set from the raw data downloaded from the following URL:

https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip

This data was collected by a group at the University of California at Irvine.  The URL of the website for their project is:

http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones


Below are the column names in the tidy data set (both the original name and the descriptive variable name I applied):

```{r echo=FALSE}
## Set some directory locations

currentwd <- getwd()

setwd("C:/Temp/data/unzipped")

topdir <- "UCI HAR Dataset"
#testloc <- paste(topdir, "/test", sep="")
#trainloc <- paste(topdir, "/train", sep="")


featNames <- function(toplevdir) {
  ## Use the original column names from the original data set (located in the "features" file.
  xcolnames <- read.table(paste(toplevdir, "/features.txt", sep=""))
  xcolnames <- xcolnames[,2]

  ## Use gsub() to replace various components of the feature names.
  xcolnames <- gsub("^f", "mean-freq-", xcolnames)
  xcolnames <- gsub("^t", "mean-time-", xcolnames)
  xcolnames <- gsub("BodyBody", "Body-", xcolnames)
  xcolnames <- gsub("Body", "Body-", xcolnames)
  xcolnames <- gsub("Gravity", "Gravity-", xcolnames)
  xcolnames <- gsub("Mag", "Magnitude", xcolnames)
  xcolnames <- gsub("mean\\(\\)", "mean", xcolnames)
  xcolnames <- gsub("std\\(\\)", "st_dev", xcolnames)
  xcolnames <- gsub("--", "-", xcolnames)
  
  return(xcolnames)
}

origcolnames <- read.table(paste(topdir, "/features.txt", sep=""))
origcolnames <- as.vector(origcolnames[,2])

#origcolnames <- as.vector(rbind("subjectId", "activityLabel", origcolnames))

#print(str(origcolnames))

xcolumnnames <- as.vector(featNames(topdir))
#xcolumnnames <- as.vector(rbind("subjectId", "activityLabel", xcolumnnames))

#print(str(xcolumnnames))

xcolumnnames <- as.data.frame(cbind(xcolumnnames, origcolnames))

names(xcolumnnames) <- c("Descriptive Variable Name", "Original Variable Name")
#print(xcolumnnames)
#print(str(xcolumnnames))
#print(dim(xcolumnnames))

meansd <- sapply(xcolumnnames[,1], function(x) any(grep(paste(c('activityId','subjectId','-mean','st_dev'),collapse='|'), x)))
xcolnms <- xcolumnnames[meansd,]

## Remove the meanFreq columnns.
meansd <- sapply(xcolnms[,1], function(x) any(!grepl('meanFreq', x)))
xcolnms <- xcolnms[meansd,]

#xcolnms <- rbind(cbind(c("subjectId","subjectId"),c("activityLabel", "activityLabel")),xcolnms)

row.names(xcolnms) <- 1:nrow(xcolnms)

print(xcolnms)
setwd(currentwd)
```





You can also embed plots, for example:

```{r}

```

