CODEBOOK FOR GETTING AND CLEANING DATA PROJECT
========================================================

The Getting and Cleaning Data Coursera Course Project involved preparing a tidy data set from the raw data downloaded from the following URL:

https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip

This data was collected by a group at the University of California at Irvine.  The URL of the website for their project is:

http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones


Below are the column names in the tidy data set (both the original name and the descriptive variable name I applied):

```{r echo=FALSE}
## Set some directory locations

currentwd <- getwd()

setwd("C:/Temp/data/unzipped")

topdir <- "UCI HAR Dataset"
#testloc <- paste(topdir, "/test", sep="")
#trainloc <- paste(topdir, "/train", sep="")


featNames <- function(toplevdir) {
  ## Use the original column names from the original data set (located in the "features" file.
  xcolnames <- read.table(paste(toplevdir, "/features.txt", sep=""), stringsAsFactors=FALSE)
  xcolnames <- xcolnames[,2]

  ## Use gsub() to replace various components of the feature names.
  xcolnames <- gsub("^f", "mean-freq-", xcolnames)
  xcolnames <- gsub("^t", "mean-time-", xcolnames)
  xcolnames <- gsub("BodyBody", "Body-", xcolnames)
  xcolnames <- gsub("Body", "Body-", xcolnames)
  xcolnames <- gsub("Gravity", "Gravity-", xcolnames)
  xcolnames <- gsub("Mag", "Magnitude", xcolnames)
  xcolnames <- gsub("mean\\(\\)", "mean", xcolnames)
  xcolnames <- gsub("std\\(\\)", "st_dev", xcolnames)
  xcolnames <- gsub("--", "-", xcolnames)
  
  return(xcolnames)
}

origcolnames <- read.table(paste(topdir, "/features.txt", sep=""),stringsAsFactors=FALSE)
origcolnames <- as.vector(origcolnames[,2])

# origcolnames <- as.vector(c("activityLabel",origcolnames))
# print(str(origcolnames))
# origcolnames <- c("subjectId",origcolnames)
# print(str(origcolnames))

#origcolnames <- as.vector(rbind("subjectId", "activityLabel", origcolnames))

#print(str(origcolnames))

xcolumnnames <- as.vector(featNames(topdir))
#xcolumnnames <- as.vector(rbind("subjectId", "activityLabel", xcolumnnames))

#xcolumnnames <- as.vector(c("activityLabel",xcolumnnames))
#print(str(xcolumnnames))
#xcolumnnames <- c("subjectId",xcolumnnames)
#print(str(xcolumnnames))

#print(str(xcolumnnames))

#xcolumnnames <- as.data.frame(cbind(xcolumnnames, origcolnames),stringsAsFactors=FALSE)
xcolumnnames <- as.data.frame(cbind(xcolumnnames, origcolnames),stringsAsFactors=FALSE)

names(xcolumnnames) <- c("Descriptive Variable Name", "Original Variable Name")
#print(xcolumnnames)
#print(str(xcolumnnames))
#print(dim(xcolumnnames))

meansd <- sapply(xcolumnnames[,1], function(x) any(grep(paste(c('activityId','subjectId','-mean','st_dev'),collapse='|'), x)))
xcolnms <- xcolumnnames[meansd,]

#print(str(xcolnms))

## Remove the meanFreq columnns.
meansd <- sapply(xcolnms[,1], function(x) any(!grepl('meanFreq', x)))
xcolnms <- xcolnms[meansd,]

#xcolnms <- rbind(as.data.frame(c("subjectId","subjectId")),as.data.frame(c("activityLabel", "activityLabel")),xcolnms)
xcolnms <- rbind(c("subjectId","subjectId"),c("activityLabel", "activityLabel"),xcolnms)

row.names(xcolnms) <- 1:nrow(xcolnms)

print(xcolnms)
setwd(currentwd)
```


A more detailed description of each variable follows:

```{r, echo=FALSE}
#subjectId
print(xcolnms[1,1],row.names=FALSE)
```

There are 30 subjects in this study overall.  This ID uniquely identifies each subject.  
There are no units for this variable.

```{r, echo=FALSE}
#activityLabel
print(xcolnms[2,1], row.names=FALSE)
```

Each subject performed multiple activities.  This descriptive label uniquely identifies the activity that was being performed by the subject when the associated measurements were taken.  
There are no units for this variable.

```{r, echo=FALSE}
#mean-time-Body-Acc-mean-X
print(xcolnms[3,1], row.names=FALSE)
```

This variable is the average of the measurements, in the time domain, of the acceleration componentof the body in the x-direction for each given activity and for each subject.  
The units for this variable are .